Automatic prostate cancer detection through \acs*{dce}-\acs*{mri} images: all you need is a good normalization

### 2016-08-19 02_methods/methods.tex
Finish to explain the graph based approach

>>>>

\begin{figure}
  \centering
  \hspace*{\fill}
  \subfigure[Patient 1]{\label{subfig:pat1}\includegraphics[width=.49\textwidth]{02_methods/figures/pat1.pdf}} \hfill
  \subfigure[Patient 2]{\label{subfig:pat2}\includegraphics[width=.49\textwidth]{02_methods/figures/pat2.pdf}}
  \hspace*{\fill}
  \caption{Illustration of the variations of the intensity \acs*{pdf} over time of two patients in a \ac{dce}-\ac{mri}.}
  \label{fig:heatmap}
\end{figure}

In this work, we proposed a method to normalized \ac{dce}-\ac{mri} prostate data, although it can be ported to any \ac{dce}-\ac{mri} sequences.
The aim of the method is to reduce the intra-patient variations that can occur during acquisition.
In \ac{t2w}-\ac{mri} prostate image, an offset and a scale factor are the two factors driving the intra-patient variation.
Therefore, these variations are corrected using a $z$-score approach assuming that the data follow a Rician distribution~\citep{lemaitre2016normalization}.
In \ac{dce}-\ac{mri}, there is additional factors characterizing such variations.
These variations can be highlighted by observing the evolution of the intensity \ac{pdf} of the \ac{dce}-\ac{mri} over time, as shown in the heatmap in Fig.\,\ref{fig:heatmap}.
These variations are:
(i) an offset of the peak value before pre-contrast,
(ii) a time offset depending of the contrast injection, and
(iii) a global scale factor related to the enhancement.
Therefore, our normalization method should attenuate all these variations and be performed globally across the different time sequence rather than for each independent sequence.

\subsubsection{Graph-based offset correction for each \ac{dce}-\ac{mri} sequence}

\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{02_methods/figures/estimator.pdf}
  \caption{Illustration of the estimator found using the shortest-path through the graph.}
  \label{fig:estimator}
\end{figure}

Before to standardize each sequence, the first step of the normalization is to cancel the intensity shift which is specific for each patient.
The intensity \ac{pdf} do not follow either a Rician or Gaussian distribution over time.
Therefore, the mean cannot be used as a potential estimate for the offset.
Additionally, this offset should be characterized by a smooth transition over time.
This problem can be solved using the graph-theory: considering the intensity \ac{pdf} over time as shown in Fig.\,\ref{fig:heatmap}, the offset is the boundary splitting the heatmap in two partitions such that it is as close as possible to the peak of the intensity \ac{pdf} (see Fig.\,\ref{fig:estimator} for an illustration).
Given the heatmap, a directed weighted graph $\mathcal{G}=(\mathcal{V}, \mathcal{E})$ is built by taking each bar of the heatmap as a node and connecting each pair of bars by an edge.
The edge weight $w_{ij}$ between two nodes $i$ and $j$ corresponding to two pixels at position $(x_i, y_i)$ and $(x_j, y_j)$, respectively, is defined as in Eq.\,\eqref{eq:weight}:

\begin{equation}
  w_{ij} = \begin{cases}
    \alpha \exp(1 - \frac{H(i)}{\max(H)})       & \text{if } x_j = x_i + 1 \text{ and } y_j = y_i \\
    (1 - \alpha) \exp(1 - \frac{H(i)}{\max(H)}) & \text{if } x_j = x_i \text{ and } y_j = y_i + 1 \\
    0                                           & \text{otherwise}
  \end{cases} ,
  \label{eq:weight}
\end{equation}

\noindent where $H$ is the heatmap, $\alpha$ is a smoothing parameter controlling the partitioning.

Therefore, the offset is estimated by finding the shortest-path to cross the graph using Dijkstra algorithm.
The entry and exiting nodes are set to be the bin with the maximum probability for the first \ac{dce}-\ac{mri} serie and the bin corresponding to the median value for the last \ac{dce}-\ac{mri} serie, respectively.
To ensure a robust estimation of the offset, the process of finding the shortest-path is iteratively repeated by shifting the data and updating the heatmap as well as the graph $\mathcal{G}$.
The procedure is stopped once that the offset found do not change.
In general, this process is not repeated more than 3 iterations.
An example of the offset found using this approach is presented in Fig.\,\ref{fig:estimator}.

\subsubsection{Estimate the dispersion of the voxel intensities}

MSE.

\subsubsection{Parametric curve alignment}

Model using 1 translation and one scale.
